---
title: "SNAP Project"
author: "Luke Wejman, Jasmine Meyer, Benjamin Auby, Bruce Garcia-Morales"
date: today
format: 
  pdf:
      toc: true
      toc-depth: 4
      shift-heading-level-by: 2
      fig-pos: "H"
      fig-cap-location: top
      geometry:
        - top=1in
        - right=.8in
        - bottom=1in
        - left=.8in
      link-citations: true
      linkcolor: blue
      include-in-header: 
        text: |
          \usepackage{fancyhdr}
          \usepackage{titling}
          \pagestyle{fancy}
          \fancyhf{}
          \renewcommand\maketitle{
            \fancyhead[C]{
              \thetitle
              \ifx \theauthor\empty  \else \ – \theauthor \fi
              \ifx \thedate\empty  \else \ – \thedate \ \fi
            }
          }
          \fancyfoot[C]{\thepage}
---

```{r}
#| echo: false
#| output: false
#| message: false

# Clear your environment
rm(list=ls())

# Install packages below if you do not have them:
# -------------------------------------------------
if (!"statnet" %in% installed.packages()) install.packages("statnet") # For fitting ERGMs
if (!"igraph" %in% installed.packages()) install.packages("igraph") # For network plotting
if (!"texreg" %in% installed.packages()) install.packages("texreg") # For printing "nicer" model output

library(statnet)
library(igraph)
library(ggplot2)
library(texreg)
library(network)
library(dplyr)

# -----------------------------
# Set the working directory
# Session > Set Working Directory > To Source File Location
# -----------------------------
list.files() # List the files in the current working directory to see if you're in the right directory

```

```{r}

SFF_data <- read.csv("copurchasing_science_fiction_fantasy.csv")
PF_data <- read.csv("copurchasing_parenting_families.csv")
CFW_data <- read.csv("copurchasing_cooking_food_wine.csv")
```

```{r}


amazon_data <- read.csv("amazon_copurchasing_clean.csv")
amazon_igraph <- graph_from_data_frame(amazon_data, directed = TRUE)
```

```{r}
combinedConnections <- rbind(SFF_data, PF_data, CFW_data)

igraph <- graph_from_data_frame(
  combinedConnections
)
igraph

```

```{r}
igraph_options(
  vertex.size = 12,
  vertex.color = 'grey',
  edge.color = 'gray80',
  edge.arrow.size = 0.1,
  vertex.label = NA
)

```

```{r}
net_layout <- layout_with_fr(igraph)
```

```{r}
plot(
  igraph,
  layout = net_layout,
  edge.color = 'black',
  vertex.label = V(igraph)$name, # Displays node/ASIN IDs
  vertex.label.cex = 0.5,
  margin = -0.001
)

```

```{r}


```

```{r}

```

# ERGM

```{r}

SFF_data <- read.csv("copurchasing_science_fiction_fantasy.csv")
PF_data <- read.csv("copurchasing_parenting_families.csv")
CFW_data <- read.csv("copurchasing_cooking_food_wine.csv")
combinedConnections <- rbind(SFF_data,PF_data, CFW_data)

# vertex set
SFF_meta <- read.csv("metadata_science_fiction_fantasy.csv")
PF_meta <- read.csv("metadata_parenting_families.csv")
CFW_meta <- read.csv("metadata_cooking_food_wine.csv")
combined_meta <- rbind(SFF_meta, PF_meta, CFW_meta)
```

```{r}
# Snowball sampling
valid_ids <- as.character(combined_meta$id)

filteredConnections <- combinedConnections %>%
  mutate(across(1:2, as.character)) %>%
  filter(.[[1]] %in% valid_ids,.[[2]] %in% valid_ids)

set.seed(42)
g <- graph_from_data_frame(filteredConnections, directed = TRUE)
g <- simplify(g,remove.multiple = TRUE,)
seeds <- sample(V(g)$name,150)

# Nodes & Edges
snowball_nodes <- ego(g,order = 20,nodes = seeds,mode = "all")
snowball_nodes <- unique(unlist(snowball_nodes))
snowball_graph <- induced_subgraph(g, vids = snowball_nodes)


# Match sampled products to metadata

product_ids <- V(snowball_graph)$name
snowball_meta <- combined_meta[match(product_ids,as.character(combined_meta$id)),]
snowball_edges <- igraph::as_data_frame(snowball_graph,what = "edges"
)

# Final sample of nodes & edges
head(snowball_meta)
head(snowball_edges)

vcount(snowball_graph)
ecount(snowball_graph)
```

### Hypotheses

***Hypothesis 1***: There will be indegree popularity effects — that is, a tendency for a small number of products to be frequently co-purchased alongside many others (as opposed to co-purchase links being spread evenly across all products).

***Hypothesis 2***: There will be outdegree effects — that is, some products will tend to "point to" (be purchased alongside) an unusually large number of other products, beyond what chance alone would predict.

***Hypothesis 3***: If product *i* is co-purchased with product *j*, it will be more likely that product *j* is also co-purchased with product *i* (reciprocity).

***Hypothesis 4***: Co-purchase relationships will tend to be transitive — that is, if product *i* is co-purchased with product *k*, and *k* is co-purchased with product *j*, then *i* will be more likely to be co-purchased with *j* as well.

***Hypothesis 5***: Products with a better (lower) sales rank will be more likely to be co-purchased with other products, as popularity drives bundling behavior.

***Hypothesis 6***: Products with a higher total number of reviews will be more likely to be sought out as co-purchase partners by other products.

***Hypothesis 7***: Products with a higher average review rating will be more likely to be co-purchased with other products, as quality signals encourage bundled buying.

```{r}

# create Amazon network
amazonEdgelist <- data.frame(
  from = match(
    as.character(snowball_edges$from),
    as.character(product_ids)
  ),
  to = match(
    as.character(snowball_edges$to),
    as.character(product_ids)
  )
)

amazonEdgelist <- unique(amazonEdgelist)
amazon_net <- network::network.initialize(
  n = length(product_ids),
  directed = TRUE
)

# Each vertex gets its actual Amazon id
network::set.vertex.attribute(
  amazon_net,
  "vertex.names",
  value = as.character(product_ids)
)
network::add.edges(
  amazon_net,
  tail = amazonEdgelist$from,
  head = amazonEdgelist$to
)

network.size(amazon_net)
network.edgecount(amazon_net)


```

```{r}
# set vertex attributes

amazon_net |> network::set.vertex.attribute(
  "salesrank",
  value = snowball_meta$salesrank
) 

amazon_net |> network::set.vertex.attribute(
  "reviews_total",
  value = snowball_meta$reviews_total
) 

amazon_net |> network::set.vertex.attribute(
  "reviews_avg_rating",
  value = snowball_meta$reviews_avg_rating
) 

amazon_net |> network::set.vertex.attribute(
  "categories_count",
  value = snowball_meta$categories_count
) 

amazon_net |> network::set.vertex.attribute(
  "product_id",
  value = snowball_meta$id
)

amazon_net |> network::set.vertex.attribute(
  "asin",
  value = snowball_meta$asin
)

amazon_net |> network::set.vertex.attribute(
  "title",
  value = snowball_meta$title
)
amazon_net
```

```{r}
# ERGM Model
start_time <- Sys.time()
ergm_Model <- ergm(amazon_net ~ edges  
               + mutual
               + gwidegree(log(2), fixed = T)  # Inverted preferential attachment (indegree)
               + gwodegree(2, fixed = T)  # Inverted preferential attachment (outdegree)
               + dgwesp(log(2), type = "OTP", fixed = T)
               
               # node attribute effects
               + nodeicov("salesrank")
               + nodeicov("reviews_total")
               + nodeicov("reviews_avg_rating")
               + nodeicov("categories_count")
              
               # Model constraints
               ,control = control.ergm(seed = 42)
               ,verbose = F
)

end_time <- Sys.time()
elapsed_time <- end_time - start_time
elapsed_time
summary(ergm_Model)

```

```{r}
# Model Diagnositcs
par(mar=c(1,1,1,1))
mcmc.diagnostics(ergm_Model, 
                 which="plots" # this command produces only the output plots
                 )
```

```{r}
#| echo: false
#| warning: false

# -------------------------------------------------------------------------------------------------
# Goodness of fit test - will display in RStudio
# Check how well the estimated model captures certain features of the observed network, for example triangles in the network.
# -------------------------------------------------------------------------------------------------

# Calculate layout (node coordinates) from the OBSERVED network
amazon_igraph <- graph_from_adjacency_matrix(as.matrix.network(amazon_net))
net_layout <- layout_with_kk(amazon_igraph)

# This first command simulates 100 networks.
# These networks, if we use sufficient burnin steps in the markov chain used to generate them,
# may be thought of as random samples from the joint probability distribution that is our fitted ERGM.
sim1 <- simulate(ergm_Model, burnin=100000, interval=100000, nsim=100, verbose=T)  # Uses the ergm model to simulate a null model

sim1_net1 <- igraph::graph.adjacency(as.matrix.network(sim1[[1]]))
igraph::plot.igraph(sim1_net1,layout=net_layout,edge.color="grey",  
                    vertex.color = 'grey',edge.arrow.size=.1)                                                       

# Plot the 10th simulated network
sim1_net10 <- igraph::graph.adjacency(as.matrix.network(sim1[[10]]))
igraph::plot.igraph(sim1_net10,layout=net_layout,edge.color="purple",  
                    vertex.color = 'grey',edge.arrow.size=.1)


```

```{r}
#| warning: false
start_time <- Sys.time()
# Model 2:
# It may take a second for this command to run.
gof1 <- gof(ergm_Model, verbose=T, burnin=1e+5, interval=1e+5, control = control.gof.ergm(nsim = 200))
# If you run below and then wouldn't see the plot, trypar(mar=c(2,2,2,2))
# dev.off()           # Clear any other plots from the plot window
plot(gof1)          # Plot the goodness of fit
# Note: This should produce five separate plots that you should look through.
#       In RStudio, scroll between the plots using the arrow buttons
gof1                # Display the goodness of fit info in the console


end_time <- Sys.time()

elapsed_time <- end_time - start_time

elapsed_time
```
