---
title: "Lab 3"
author: "Luke Wejman"
date: today
format: 
  pdf:
      toc: true
      toc-depth: 4
      shift-heading-level-by: 2
      fig-pos: "H"
      fig-cap-location: top
      geometry:
        - top=1in
        - right=.8in
        - bottom=1in
        - left=.8in
      link-citations: true
      linkcolor: blue
      include-in-header: 
        text: |
          \usepackage{fancyhdr}
          \usepackage{titling}
          \pagestyle{fancy}
          \fancyhf{}
          \renewcommand\maketitle{
            \fancyhead[C]{
              \thetitle
              \ifx \theauthor\empty  \else \ – \theauthor \fi
              \ifx \thedate\empty  \else \ – \thedate \ \fi
            }
          }
          \fancyfoot[C]{\thepage}
---

```{r}
#| echo: false
#| output: false
#| message: false

# Clear your environment
rm(list=ls())

# Install packages below if you do not have them:
# -------------------------------------------------
if (!"statnet" %in% installed.packages()) install.packages("statnet") # For fitting ERGMs
if (!"igraph" %in% installed.packages()) install.packages("igraph") # For network plotting
if (!"texreg" %in% installed.packages()) install.packages("texreg") # For printing "nicer" model output

library(statnet)
library(igraph)
library(ggplot2)
library(texreg)

# -----------------------------
# Set the working directory
# Session > Set Working Directory > To Source File Location
# -----------------------------
list.files() # List the files in the current working directory to see if you're in the right directory

```

# Lab 3: Exponential Random Graph Models (ERGMs)

In this lab, we will be testing hypotheses about a network's structure using exponential random graph modeling (ERGM) techniques using the **statnet** package in R. The **statnet** package provides a comprehensive framework for ERGM-based network modeling, including tools for model estimation, model evaluation, model-based network simulation, and network visualization. More documentation about **statnet** is available here: [http://csde.washington.edu/statnet**/**](http://csde.washington.edu/statnet/). More information about ERGM itself can be found in this supplemental lecture video: <https://bit.ly/ERGM_Explanations>. The slides presented in this video are also available on Canvas.

The data used in this lab is from a people analytics project. People analytics refers to the use of HR data to understand behavior in organizations. We will analyze data about the social relations among employees in a small Chinese company. We are interested in understanding who employees go to for advice.

Our dependent variable examines survey responses to the question:

**"List up to 5 employees who you rely on the most for help or advice at work.**"

These responses form a directed network where there is a tie directed from the survey respondent to the person they go to for advice. We provide this data to you in the format of an edgelist, *adviceEdgelist.csv*, in which all 66 employees have been assigned an id ranging from 1 to 66.

Our objective will be to determine what factors influence who someone goes to for advice. We will create a model of this Advice network based on data that is available to the company.

This includes typical HR information, on node attributes, available to them:

● *departmentNode.csv* contains text on which department each employee belongs to.

● *officeNode.csv* contains nodes office locations: 1 for the main office, 0 for the secondary.

● *tenureNode.csv* contains the tenure at the company of each node, in years.

● *leaderNode.csv* contains an indicator (binary) for whether each node is a department lead.

● *femaleNode.csv* contains an indicator variable for sex. 1 for female, 0 for male.

What is unique about this dataset is that we will also look at digital trace data available to the employer. They use an Enterprise Social Media (ESM) platform to allow communication between employees. Enterprise Social Media are social networks developed for work-related use, for example Microsoft Teams or Slack. We have collected data on the number of direct messages employees have sent to one another over roughly six weeks' time. This is shared in the form of an edgelist in *messageEdgelist.csv*.

## **Hypotheses:**

We are going to test the following hypotheses in this lab.

***Hypothesis 1***: There will be indegree popularity effects -- That is, a tendency for a small number of employees to be sought out for advice from many others (as opposed to advice seeking behaviors being spread evenly amongst all employees).

***Hypothesis 2***: Individuals will be more likely to report going to people in their own department, as opposed to other departments, for advice.

***Hypothesis 3***: There will be homophily based upon the sex of individuals, in terms of who employees go to for advice.

***Hypothesis 4***: Employees who message someone more frequently on ESM will be more likely to report going to that person for advice.

***Hypothesis 5***: If an employee *i* goes to another employee *j* for advice, it will be more likely that *j* also goes to employee *i* for advice.

***Hypothesis 6***: Employees who work in the main office will be more likely to go to others for advice than employees from the secondary office.

***Hypothesis 7***: Employees who work in the main office will be more likely to be sought after for advice than employees from the secondary office.

***Hypothesis 8***: Advice seeking relationships tend to be transitive - That is, if individual *i* goes to an individual *k* for advice, and *k* goes to an individual *j* for advice*,* then *i* is more likely to go to *j* for advice as well.

These are the hypotheses that we are interested in testing. However, you will notice that the code for ERGMs you have been provided contains more ERGM terms than just the ones corresponding to the above hypotheses. This is normal -- in testing the hypotheses of interest, we may still want to include additional model terms to act as controls for other factors.

::: {.callout-note style="color: purple"}
Your job is to run the code and, most critically, interpret the output. Test the above hypotheses and prepare the report on your work. Include a copy of the relevant output (model parameters, plots) in your report.
:::

# Part I: Building and Visualizing the Networks (20 pts)

The analysis will use 3 types of files: the "adviceEdgelist.csv" as the *base* *network* file (the ties the model is predicting), "messageEdgelist.csv" as the *covariate network* file, and the remaining csv files as data on *node attributes*. Download all these files into your working directory and load the network and attribute data.

```{r}
#| echo: false
#| output: false
#| message: false

# First we will load in the adviceEdgelist.csv file, which contains our dependent
# variable. These are responses to the question:
# “List up to 5 employees who you rely on the most for help or advice at work.” 
# Note that participants were limited to selecting at most five respondents.
adviceEdgelist <- read.csv("adviceEdgelist.csv")
# View the first rows of the edgelist to make sure it imported correctly:
head(adviceEdgelist)
# Convert the edgelist to a network object in statnet format:
advice <- as.network.matrix(adviceEdgelist, matrix.type = "edgelist") 
# View a summary of the network object
advice
# Independent variables:
# Load node attributes, and store them in the advice network object we have created
advice |> network::set.vertex.attribute("department", 
                          value = read.csv("departmentNode.csv")$department) # Categorical variable for department

advice |> network::set.vertex.attribute("leader",read.csv("leaderNode.csv")$leader) # Indicator variable for department leader
advice |> network::set.vertex.attribute("tenure",read.csv("tenureNode.csv")$tenure) # Years tenure
advice |> network::set.vertex.attribute("office",read.csv("officeNode.csv")$office) # Indicator variable for whether they are located in the main or secondary office
advice |> network::set.vertex.attribute("female",read.csv("femaleNode.csv")$female) # Indicator variable for female vs. male
advice # These five variables should now be listed as vertex attributes when viewing the summary of the network
```

```{r}
#| echo: false
#| output: false

# Double-check the values for all of the node variables
network::get.vertex.attribute(advice,"department")
network::get.vertex.attribute(advice,"leader")
network::get.vertex.attribute(advice,"tenure")
network::get.vertex.attribute(advice,"office")
network::get.vertex.attribute(advice,"female")
```

```{r}
#| echo: false
#| output: false

# Finally, we will import data on the counts of direct messages sent between two employees
# The file "messageEdgelist.csv" contains a messaging edgelist, that we will convert to a matrix
# Statnet allows adjacency matrices with valued ties to be used as predictors (edge covariates) in ERGMs
messageEdgelist <- read.csv("messageEdgelist.csv")
head(messageEdgelist) # Check the first five rows of the edgelist. Third column is the message count
messages <- matrix(nrow = 66, ncol = 66) # Number of Direct messages sent from i to j
for (i in 1:nrow(messageEdgelist)) { # Read from edgelist
  messages[messageEdgelist$SenderId[i], messageEdgelist$ReceiverId[i] ] <- as.numeric(messageEdgelist$MessagesSent[i])
}
for (i in 1:66) { # Remove self-ties (messages sent to self)
  messages[i,i] <- as.numeric(0)
}
hundreds_messages <- messages / 100 # Change weights to represent hundred of messages sent
                                   # This will make viewing/interpreting ergm coefficients easier

# ---------------------------------------------------------------------------------------
# Basic descriptive information
# ---------------------------------------------------------------------------------------
summary(advice, print.adj = FALSE)           # summarize the advice From You network
```

## 1. Describe the data that you are analyzing. **(5 pts)**

#### What network(s) are you analyzing? What edge or vertex attributes are stored? Are there any limitations to the data or assumptions based on how it was collected?

Answer: The data that we are analyzing is a group of networks that is descriptive of a workplace. A vertex refer to essentially an employee in said workplace, which can also be sorted or categorically grouped by department, sex, main office, etc. The edges represent a directional relationship based on advice seeking, i.e. employee can seek advice from employee 2, so there would be an edge from node 1 to 2. It terms of how the data was collected, we can assume the majority of it was company records, such as sex and departmental information, but things involving advice seeking could have been sourced by asking the employees directly, and there is the potential of getting differing result per person, which can lead to inconsistent data.

## 2. Plot the base (Advice) network and interpret the visualization. **(5 pts)**

#### Explain whether this plot *seems*, at a glance, to match what you would expect to see if hypothesis 1 were true. Think of this as just a basic descriptive check -- we will perform a more rigorous statistical test in part II of the lab.

```{r}
#| echo: false
#| message: false 
# Ignore messages on any objects that are masked

# Set default plot options
igraph_options(vertex.size = 12, vertex.color = 'grey', # vertex.size changes the size of nodes; vertex.color changes the color of nodes
               edge.color='gray80', edge.arrow.size=.1, # edge.color changes the color of ties; edge.arrow.size changes the size of tie arrow heads
               vertex.label = NA)                       # vertex.label = NA specifies not to display vertex labels in the plot

# Plot the Advice network
advice_igraph <- graph_from_adjacency_matrix(as.matrix.network(advice)) # make an igraph network object from statnet network object
advice_igraph <- set_vertex_attr(advice_igraph,"female",value = read.csv("femaleNode.csv")$female)
net_layout <- layout_with_kk(advice_igraph) # Calculates and stores a spring-embedded layout

# We will re-use this same layout for each plot, so nodes are always in the same place
plot(advice_igraph, layout=net_layout, edge.color='black', 
     vertex.label = V(advice_igraph),
     vertex.label.cex = 0.5,
     margin=0
     )
```
Answer: If I were to assume hypothesis 1 was correct, this graph above does look pretty accurate. Hypothesis 1 argues that there would a select group of employees that other employees would rely on more than others, and the distribution would be less even. We see this is the graph; there are several nodes in the center that have very high degrees, and as you go farther out from the center, the degree lowers per node, to the point were some outer node only have a 1-4 connections.


## 3. Plot the base network with the nodes now colored based on sex and interpret the new visualization. **(5 pts)**

#### Explain whether this plot seems to match what you would expect to see if hypothesis 3 were true.

```{r}
#| echo: false
#| message: false 

# Plot the Advice network with node coloring based on sex
V(advice_igraph)$color = ifelse (V(advice_igraph)$female ==1, " orange ", " grey ")
plot(advice_igraph, 
     layout=net_layout, 
     edge.color='black', 
     vertex.label = V(advice_igraph),
     vertex.label.cex = 0.5,
     margin=0)
```
Answer: Assuming that hypothesis 3 is true, the graph generally meets my expectations. There are definitely groups of nodes that are all the same sex, and generally point to eachother, but there are noticeable outliers as well, for example node 47 (female) only points to node 52 (male), or node 3 (female) points to nodes 52 and 11 (male) as well as node 64 (female). While there are definitely features that would support this hypothesis, I think a more in depth analysis is necessary. 


## 4. Plot the covariate network and interpret this visualization. **(5 pts)**

#### Comparing the network plots, explain whether this plot seems to match what you would expect to see if hypothesis 4 were true.

Answer: This plot does not seem to match hypothesis 4. There definitely is overlap between this network and the first network we graphed. However, there are noticably more connections in this graph compared the other graphs, indicating that not only is there more communication on over the ESM system, but also just because there is communication between two people (nodes) in this setting, does not mean it would happen then.
```{r}
#| echo: false
#| message: false 

# Plot the network of who messages whom
# One unit of weight is 100 messages
messages_igraph <- graph_from_adjacency_matrix(messages, weighted = TRUE) ## weighted = TRUE creates edge weight in the igraph object
## below because it is added not only edge weight, but also changed the transparency of edges, it will take more time to plot
## also, if you use zoom in RStudio, it may take about 1 minute until you see the plot.
plot(messages_igraph, 
     layout=net_layout, 
     edge.color = adjustcolor('blue',alpha=.2), 
     vertex.label = V(advice_igraph), 
     vertex.label.cex = 0.5,
     margin=0,
     edge.width=log(E(messages_igraph)$weight), 
     edge.arrow.width =.2)
```


# Part II: Model Estimation (50 pts)

## 1. Build two ERGM models to test the hypotheses using the different network statistics described below **(10 pts)**

```{r}
#| echo: false
#| message: false 
#| output: false

# R vignette for more details: https://cran.r-project.org/web/packages/ergm/ergm.pdf

options(ergm.loglik.warn_dyads=FALSE) #Whether or not a warning should be issued when sample space constraints render the observed number of dyads ill-defined

# Ergm Terms are statistics: They are some deterministic function of the ties, node attributes, and edge covariates of a network.
help("ergm-terms",package = "ergm") # Documentation that contains definitions for all of the terms we are using
                                    # ex. what does "mutual" test and how is it calculated
# We will use the ergm-terms to perform hypothesis testing using ERGMs
# But we can note that any of the ERGM terms can also be examined directly for your observed network, by creating a formula in R

# Look at Endogenous statistics: terms based on only ties in the advice network
summary(advice ~ edges)                     # Number of edges (ties)
summary(advice ~ mutual)                    # Number of pairs of reciprocated ties
summary(advice ~ odegree(0:5))              # Outdegree distribution. (# of nodes with outdegree of 0, # nodes outdegree of 1, etc.)
                                           # Remember, respondents could nominate at most five employees in our survey
summary(advice ~ idegree(0:65))             # Indegree distribution.
summary(advice ~ gwodegree(log(2),fixed=T)) # One parameter summarizing outdegree distribution - tendency against outdegree hubs
summary(advice ~ gwidegree(log(2),fixed=T)) # One parameters summarizing indegree distribution - tendency against indegree hubs
summary(advice ~ desp(1:5))                 # Pairs of nodes with one shared partner, two shared partners, etc.
summary(advice ~ dgwesp(log(2),fixed = T))  # One parameter summarizing 

# Look at Exogenous statistics: terms based on advice ties AND other ties / node attributes
summary(advice ~ nodeicov("office"))             # Ties directed towards employees at the main office (as opposed to secondary office)
summary(advice ~ nodeocov("office"))             # Ties originating from employees at the main office (as opposed to secondary office)
summary(advice ~ nodematch("female"))            # Number of ties between people of the same sex
summary(advice ~ nodematch("department"))        # Number of ties between people working in the same department
summary(advice ~ nodemix("leader",levels2=NULL)) # Number of ties between different combinations of leaders(1) and non-leaders(0)
summary(advice ~ diff("tenure"))                 # Total difference in tenure: sum of (sender's tenure - receivers-tenure) for all ties
summary(advice ~ edgecov(hundreds_messages))     # Total messages sent: sum of (messages sent from sender to receiver)/100 for all Advice ties
                                                # e.g., a total of 5669 messages were sent from employees to those they go to for advice during the observed period

# The above are statistics - counts of these patterns for our networks
# What fitting the ERGM model will tell is whether these counts are relatively high/low
# in comparison to what we would expect based on random chance, controlling for the other effects in our model.
# This type of analysis can be helpful for understanding your network, as well as troubleshooting issues with ERGM regression
```

#### Look at the endogenous and exogenous summary statistics. Based on these values, how many pairs of advice-seeking ties were reciprocated? How many advice-seeking ties originated from the main office?

Answer: There are 12 reciprocated advice seeking ties, and 173 advice seeking ties originated from the main office. 

::: {.callout-note style="color: purple"}
Fit model 1 (simple model) and model 2 (complex model) using the terms already specified in the R script provided to you. We will start with the simple model.
:::

#### What model constraints were defined in both models and why?

Answer: the constraints are the bounded outdegree constraints (bd(maxout = 5)). The reason we have it is because it is used to reflect the realistic limits on how many people an employee can seek advice from, and to prevent unrealistic networks from being generated. 

```{r}
#| echo: false
#| message: false

# The following commands do model estimation for ERGMs.
# This may take a second. Text will print in-console to update you on progress in model estimation.
model1 <- ergm(advice ~ edges                 # This is  a tendency towards a greater number of advice ties existing. Based on a statistic counting the number of ties.
               # Structural patterns
               + mutual                      # This is a tendency towards reciprocity for the advice ties. Based on a statistic counting the number of reciprocated ties.
               + edgecov(hundreds_messages)  # This is the effect of every 100 messages sent from i->j on likelihood of an advice tie. Based on a weighted sum of advice ties x 100s of messages sent
               + nodemix("leader",base = 3)
               # Model constraints
               , constraints =~ bd(maxout=5) # This constraint enforces the maximum outdegree is 5
               , control = control.ergm(seed = 42)
               ,verbose = F
)
summary(model1) 

## Convert a log-odds (e.g., -0.68122) ratio to an odds ratio
# exp(-0.68122)
```

::: {.callout-note style="color: purple"}
Now we will fit a more complex model.
:::

#### What parameter was included in the simple model but not in the complex model?

Answer: edges

```{r}
#| echo: false
#| message: false ## make it true, if you want to see cool messages about Convergence process iterations
#| warning: false
model2 <- ergm(advice ~  # This model will be slower to estimate than model 1
                        # Expect roughly 2-7 minutes. If it gets stuck for longer than that, try hitting "stop" and re-running it
               # Structural patterns
               # edges
               mutual
               + gwidegree(log(2), fixed = T)                 # Inverted preferential attachment (indegree)
               + gwodegree(2, fixed = T, cutoff = 5)              # Inverted preferential attachment (outdegree)
               + dgwesp(log(2), type = "OTP", fixed = T, cutoff =5)    # A modified version of Outgoing Two Path(i->j + i->k->j) structures. Geometrically weighted version of transitivity
               # Node attribute effects
               + nodematch("female")                                   # Homophily on a categorical variable 
               + nodemix("leader", base = 3)                           # Mixing matrix of all different combinations of node attributes (ex. A -> A ties, A-> B ties, B -> A ties, B -> B ties). 
               + nodematch("department") 
               + nodeicov("office")                                    # Covariance between in-degree of nodes and attributes of nodes
               + nodeocov("office")                                    # Covariance between out-degree of nodes and attributes of nodes
               + diff("tenure")                                        # Difference is computed as (tenure_i - tenure_j) i: sending node, j: receiving node
               + edgecov(hundreds_messages)                            # Covariance between edges of two networks (predictor can be continous)
               # Constraints on network
               , constraints =~ bd(maxout=5)                           # This constraint enforces the maximum outdegree is 5
               # Control settings for MCMC-MLE algorithm
               , control = control.ergm(MCMC.effectiveSize = 50, 
                                        seed = 42)
) 
summary(model2) 
```

## 2. For each of the eight hypotheses, interpret the results from your models and state whether that hypothesis was supported. **(40 pts)**

#### To determine whether a hypothesis is supported, look at whether there is a p-value \< 0.05 and the directionality (positive/negative) of the effect. When you interpret the results, convert the model coefficients, which are given by R as conditional log-odds, into odds ratios.

```{r}
#| echo: false

# Easy side-by-side model comparison:
texreg::screenreg(list("model1"=model1,"model2"=model2))
```

### Hypothesis 1


Answer: 
p-threshold < 0.001

-2.44/0.35 = -7.041, exp(-2.436) = 0.087. p value = 0.0001

Conclusion: Hypothesis Support, p is less that 0.001, 

### Hypothesis 2

Answer: 

p-threshold < 0.001

2.12 / 0.18346 = 11.541, 

exp(2.12) = 8.31

p value = 0.0001

Conclusion: employees are 8.31 times more likely of seeking advice from someone within their own department. Additionally the p-value is smaller than the threshold, therefore, the Hypothesis is support.



### Hypothesis 3

Answer: 

p-value threshold < 0.05

0.20605/0.14123 = 1.459

exp(0.20605) = 1.23

p-value = 0.14458

P = 0.14458 > 0.05

Conclusion: Hypothesis not support, because while the odds for seeking advice from the same gender is 1.23, the p value is larger than 0.05, which indicates the a failure of the null hypothesis.


### Hypothesis 4

Answer: 

p-value threshold < 0.001

0.408/0.0933 = 4.378

exp(0.408) = 1.50

p-value = 0.0001

Conclusion: Hypothesis supported, because the p-value is smaller than 0.001, and the odds of i seeking advice from j after 100 additional messages, it increases by 1.50 times.

### Hypothesis 5

Answer: 

p-value threshold < 0.01

-0.42963/0.49491 = -0.868

exp(-0.42963) = 0.65

p-value =  0.38534.

Conclusion: Hypothesis not supported, because the p-value  0.38534 > 0.01, thus failing to meet the threshold. Additionally, the odds are 0.65, further indicating that the probability of this being true is not very good.


### Hypothesis 6

Answer: 

p-value threshold < 0.05

0.01114/0.19181 = 0.058

exp(0.01114) = 1.01

p-value = 0.95367

Conclusion: Hypothesis not supported, The odds ration is nearly no effect with 1.01, and the p-value is well beyond the threshold: 0.95367 > 0.05. Therefore, not supported.

### Hypothesis 7

Answer: 

p-value threshold < 0.05

-0.23897/0.14316 = -1.669 = z

exp(-0.23897) = 0.79

p-value = 0.09506

Conclusion: Hypothesis not supported, the p-value threshold is not correct, 0.09506 > 0.05. Additionally, the odds are 0.79, further indicating that the probability of this being true is not very good.


### Hypothesis 8

Answer: 

p-threshold < 0.001

0.64546/0.09686 = 6.664 = z

odds ratio: exp(0.64546) = 1.91

p-value = 0.0001.

Conclusion: Hypothesis support, because the p-value is less than than 0.001, and the odds are high, with 1.91, indicating a direct relationship between i --> j. 



# Part III: Model Diagnostics (30 pts)

You can judge convergence of the MCMC process in the models using the *mcmc.diagnostics()* function. The function will plot the change of model statistics during the last iteration of the MCMC estimation procedure. For each model statistic, the left-hand side plot gives the change of the statistic with iterations, and the right-hand side plot is a histogram of the statistic values. Both are normalized, so the observed values are located at 0.

## 1. Attach the model diagnostics for model 1 and 2 in your report and interpret the plots. Has the MCMC process converged to a desired state? **(10 pts)**

### MCMC Model 1

```{r}
#| echo: false

par(mar=c(1,1,1,1))
mcmc.diagnostics(model1, 
                 which="plots" # this command produces only the output plots
                 # if you are curious about other mcmc diagnostics you can uncomment the line above.
                 )
```
Answer: For model 1, all six of the pairs of graphs look like convergence has been achieved except for edgecov.hundreds of messages, we would expect there to be a need normally straight line, and a standard bell curve. Instead we get a frequency line that is slight skewed upward, and has a humpt in the middle, the result are noticeable in the second graph, where we have a bi modal curve forming.


### MCMC Model 2

```{r}
#| echo: false

par(mar=c(1,1,1,1))
mcmc.diagnostics(model2, 
                 which="plots" # this command produces only the output plots
                 # if you are curious about other mcmc diagnostics you can uncomment the line above.
                 )
```
Answer: For model 2 it appears that every graph pair converged apart from the following: density of gwodeg.fixed.2, since the curve is skewed to the right, the nodematch.female sicne its curve is also skewed to the right, the density of mix.leader.1.0, which is skewed to the left. these are indications of convergence not happening.edgecov.hundreds_messages also has a slightly skewed distribution, but it is very minor and difficult to tell. 



## 2. Perform Goodness of Fit test to check how well the estimated model captures certain statistical features of the observed network for both model 1 and 2. **(10 pts)**

a.  To do so, simulate many networks from the estimated model and extract 100 samples from the simulation process. Please note, this may take 2 minutes or more to compute.




#### Compare the plots generated. Identify which network visualizations were generated from model 1 vs. model 2 and describe any similarities / differences you can see.

```{r}
#| echo: false
#| warning: false

# -------------------------------------------------------------------------------------------------
# Goodness of fit test - will display in RStudio
# Check how well the estimated model captures certain features of the observed network, for example triangles in the network.
# -------------------------------------------------------------------------------------------------
# Look at networks simulated according to model 1
# This first command simulates 100 networks.
# These networks, if we use sufficient burnin steps in the markov chain used to generate them,
# may be thought of as random samples from the joint probability distribution that is our fitted ERGM.
sim1 <- simulate(model1, burnin=100000, interval=100000, nsim=100, verbose=T)  # Uses the ergm model to simulate a null model
# Plot the first of the simulated networks
sim1_net1 <- igraph::graph.adjacency(as.matrix.network(sim1[[1]]))
igraph::plot.igraph(sim1_net1,layout=net_layout,edge.color="brown",  
                    vertex.color = 'grey',edge.arrow.size=.1)

# Plot the 10th simulated network
sim1_net10 <- igraph::graph.adjacency(as.matrix.network(sim1[[10]]))
igraph::plot.igraph(sim1_net10,layout=net_layout,edge.color="red",  
                    vertex.color = 'grey',edge.arrow.size=.1)


# Repeat, now looking at networks simulated according to model 2
sim2 <- simulate(model2, burnin=100000, interval=100000, nsim=100, verbose=T)  # Uses the ergm model to simulate a null model
# Plot the first of the simulated networks
sim2_net1 <- igraph::graph.adjacency(as.matrix.network(sim2[[1]]))
igraph::plot.igraph(sim2_net1,layout=net_layout,edge.color="grey",  
                    vertex.color = 'grey',edge.arrow.size=.1)                                                               
# Plot the 10th simulated network
sim2_net10 <- igraph::graph.adjacency(as.matrix.network(sim2[[10]]))
igraph::plot.igraph(sim2_net10,layout=net_layout,edge.color="purple",  
                    vertex.color = 'grey',edge.arrow.size=.1)
```

Answer: Brown and red were generated from model 1, while grey and purple from model 2. In terms of similarities, it appears that nodes in the center have a lot higher degrees compared those on the outside, and generally all four graphs appear to be one interconnected system. That being said, between each of the four graphs, the connections/edges are different. 



b.  Extract the number of triangles from each of the 100 samples.

#### How many triangles were observed in the advice seeking network? What was the mean value of triangles generated by model 1? What about in model 2?

```{r}
#| echo: false

# Observed network:
advice.tri <- summary(advice ~ triangle)                                  # Stores the number of observed triangles
advice.tri

print("Model 1")
model1.tridist <- sapply(1:100, function(x) summary(sim1[[x]] ~triangle)) # Extracts the triangle data from the simulated networks
c(obs=advice.tri,mean=mean(model1.tridist),sd=sd(model1.tridist),
  tstat=abs(mean(model1.tridist)-advice.tri)/sd(model1.tridist))

print("Model 2")
model2.tridist <- sapply(1:100, function(x) summary(sim2[[x]] ~triangle)) # Extracts the triangle data from the simulated networks
c(obs=advice.tri,mean=mean(model2.tridist),sd=sd(model2.tridist),
  tstat=abs(mean(model2.tridist)-advice.tri)/sd(model2.tridist))
```
Answer: Model 1: 153 approx. Model 2: 330 approx.


c.  Compare the distribution of triangles in the sampled networks with the observed network by generating a histogram of the triangles. 


#### Interpret your result. Are either of the models good in terms of triangle measure?

```{r}
#| echo: false

# Model 1:
#Plots that triangle distribution as a histogram
model1.tridist |> 
  as.data.frame() |> 
  ggplot(aes(x = model1.tridist)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  geom_boxplot(fill = "grey")+
  geom_vline(xintercept = advice.tri, color = "red", linewidth = 1.5) +
  geom_text(aes(x = advice.tri, y = 10, label = paste("Observed:", advice.tri)), color = "red", vjust = -1, angle = 90) +
  labs(title = "Distribution of Triangles in Model 1",
       x = "Number of Triangles",
       y = "Frequency") +
  theme_minimal()


# Model 2:
hist(model2.tridist,xlim=c(0,1000),breaks=10)                             # Plots that triangle distribution as a histogram, change xlim to change the x-axis range if necessary
arrows(advice.tri,20, advice.tri, 0.5, col="red", lwd=3)   
model2.tridist |> 
  as.data.frame() |> 
  ggplot(aes(x = model2.tridist)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  geom_boxplot(fill = "grey")+
  geom_vline(xintercept = advice.tri, color = "red", linewidth = 1.5) +
  geom_text(aes(x = advice.tri, y = 10, label = paste("Observed:", advice.tri)), color = "red", vjust = -1, angle = 90) +
  labs(title = "Distribution of Triangles in Model 2",
       x = "Number of Triangles",
       y = "Frequency") +
  theme_minimal()
```
Answer: For this, it would appear that model 1 does not produce the observed amount, and underperforms quite a bit. On the other hand, model 2 appears to be much more accurate, with 329 being close to the center of the histogram. 


## 3. Perform a goodness-of-fit evaluation process for a variety of other network statistics just for model 2 **(10 pts)**

*for example, degree distribution, distribution of edgewise shared partners, and the distribution of geodesics* Simulate networks, compile statistics for these simulations as well as the observed network, and calculate p-values of all of the aforementioned values to evaluate the correspondence between the networks simulated by the model and the observed network. Report the p-values for the simulation and interpret them.

::: {.callout-note style="color: purple"}
Below we provide sample code to perform a goodness of fit evaluation for model 1. Modify the code as necessary to perform the goodness of fit test for model 2.

Note: Your gof graph should have 5 subplots corresponding to the 5 parameters of your model. The dark black line represents the data for the observed network. The boxplots represent the distribution of corresponding degrees across the simulated networks, and the soft lines are the 95% confidence intervals. In general, for configurations in the model, the fit is considered good if │t│≤ 0.1. For configurations not included in the model, the fit is considered good if 0.1\<│t│≤ 1, and not extreme if 1\<│t│≤ 2. If │t│\> 2 the fit is bad.
:::

### Model 1 GOF

```{r}
#| warning: false
# -------------------------------------------------------------------------------------------------
# Test the goodness of fit of the model
# Compiles statistics for these simulations as well as the observed network, and calculates p-values 
# -------------------------------------------------------------------------------------------------

# Model 1:
# It may take a second for this command to run.
gof1 <- gof(model1, verbose=T, burnin=1e+5, interval=1e+5, control = control.gof.ergm(nsim = 200))
# If you run below and then wouldn't see the plot, trypar(mar=c(2,2,2,2))
# dev.off()           # Clear any other plots from the plot window
plot(gof1)          # Plot the goodness of fit
                    # Note: This should produce five separate plots that you should look through.
                    #       In RStudio, scroll between the plots using the arrow buttons
gof1                # Display the goodness of fit info in the console
```
Answer: 

Model Statistics: This gof model looks good as not only are the means very close to the observations, but the p-values are all quite large, which indicates it agrees with out null hypothesis, and the network fits well in the model statistics distribution.

Minimum Geodesic Distance: This gof model doesn't look particularly good, as about a third or so of the p-values are below 0.05, and the corresponding mean values are off from the observed. That being said there is still a fair amount of these that are accurate or very close the the observation, specifically 1, 9-18. Overall this model does not fit well within the Minimum Geodesic Distance distribution.

out-degree: This gof model isn't particularly strong either, 1/3 of the values (4 and 5) have p-values below 0.05, and their means are very different as well. That being said, there are still the remaining values above 0.05, but some of their p-values aren't too high, with values such as 0.15 and 0.26. Therefore this network does fit well within the out-degree distribution.

in-degree: This model overall looks to be very accurate, there are a handle of p-values that are below the 0.05 threshold, specifically 8/37, but there are several very high p-values, including many 1.00s. Therefore this network fits decently in in-degree distrubtion.

edgewise shared partner: Looking at the p-values, this networks fits very poorly int he edgwise shared partner metric, 3 of the 5 p-values are 0.00, and the other two are 0.14, meaning it ranks very low. 

### Model 2 GOF

```{r}
#| warning: false
# -------------------------------------------------------------------------------------------------
# Test the goodness of fit of the model
# Compiles statistics for these simulations as well as the observed network, and calculates p-values 
# -------------------------------------------------------------------------------------------------

# Model 1:
# It may take a second for this command to run.
gof2 <- gof(model2, verbose=T, burnin=1e+5, interval=1e+5, control = control.gof.ergm(nsim = 200))
# If you run below and then wouldn't see the plot, trypar(mar=c(2,2,2,2))
# dev.off()           # Clear any other plots from the plot window
plot(gof1)          # Plot the goodness of fit
                    # Note: This should produce five separate plots that you should look through.
                    #       In RStudio, scroll between the plots using the arrow buttons
gof2                # Display the goodness of fit info in the console
```
Answer:

Model-statistics: This network fits very well in model 2's model statistics distribution, every p-value is far above 0.05, many of them being close to or 1.00, indicating a very close to correct distribution. 

minimum geodesic distance: This network isn't a perfect it for the minimum geodesic distance distribution, but it is a pretty good one to say the least. Every p-value is above 0.05, the closest one is 0.08, but the majority of values are above 0.60, so while not perfect, or as good a model statistics, it is still relatively strong.

out-degree: This network is an ok fit for the out-degree. Only one p-value of the 6 if below 0.05, that being 0.04. There is also another p-value that is still very close to that threshold, it being 0.06, and the other p-values are remarkably high either, but there is not an overwhelming majority of p-values below the threshold or close to it. Therefore the distribution is relatively ok. 

in-degree: The in-degree network distribution generally fits our network, there are only two values that are equal to or less than 0.05. So while not all the p-values meet the threshold, the majority of them do, and there are many that are quite big, including several 1.00, meaning the in-degree distribution is generally strong in the network.

edgewise shared partner: Overall this network distribution of the edgewise shared partner fits ok. Of the 5, you have really good fits (0, 1, 3), but 2 is not as good of a fit (still above the threshold thought), and the 4 is below the threshold. So while so aspects of the distribution are good, others are not, so our network partially fits with the edgewise shared partner. 




## Disclose AI Use

I used no AI to generate work for this lab.

------------------------------------------------------------------------

Check your submission for grammar - points may be deducted for lack of clarity.

Click 'Render' button at the top of the screen, or press cmd + shift + k. Note. It might take some time for you computer to render this document as a PDF, since it will be running all code chunks.

Deliverables to submit on Canvas:

1.  Your report as a .pdf file
2.  Your code as a .qmd file
3.  Your data as a .RData file

```{r}
save.image(file = "lab3.RData")
```

Please upload each file separately -- do not upload as a zip file! *(Please)*

### Endogenous Effects (Effects of the ties being predicted on other predicted ties)

● ***edges***: number of edges in the network

● ***mutual***: number of reciprocal edges in the network

● ***gwidegree***: Geometrically Weighted Indegree. This term measures a tendency *against* indegree preferential attachment. (Negative coefficients show indegree preferential attachment -- Incoming ties are more likely to be directed towards nodes that already have other incoming ties.)

● ***gwodegree***: Directed Geometrically Weighted Outdegree. This term measures a tendency *against* outdegree preferential attachment. (Negative coefficients show outdegree preferential attachment -- Outgoing ties are more likely to originate from nodes that already have other outgoing ties)

● ***dgwesp***, of type "OTP": *Directed Geometrically Weighted Edgewise Shared Partners* Number of edges that belong to certain types of triangles. "Edgewise" refers to the fact that we require a tie to exist between nodes i and j, and then measure the number of "shared partners" between them. Shared partners are nodes that have a certain relationship between i and j. In this case, we are looking at the Outgoing Two Path ("OTP") relationships. This is one way to operationalize transitivity. The "geometrically weighted" refers to the fact that we will use a weight parameter,, to add diminishing returns to the number of shared partners (i.e., the second shared partner between two nodes will have less effect on the likelihood of a network than the first shared partner, the third will have even less of an effect, and so on).

Yes, geometrically weighted terms (gwidegree, gwodegree, dgwesp) are very complicated. Essentially, the "geometric weighted" part is saying that effects on network probability have diminishing returns for nodes as degree or the number of shared partners gets higher and higher. This helps avoid model fits where all the ties are directed towards one node. For the purposes of this class, you can ignore the technical details and just focus on interpreting them in terms of "preferential attachment" or "transitivity" effects.

### **Exogenous Effects (Effects of node attributes or variables outside the predicted ties)**

● ***nodeicov***: covariance between in-degree of nodes and attributes of nodes

● ***nodeocov***: covariance between out-degree of nodes and attributes of nodes

● ***diff***: differences between nodes on some numeric attribute (ex. tenure, age). The way we have it specified in the code, diff scores are is calculated as the attribute value of the sending node (attb~i~) minus value of the receiving node (attb~j~). (Heterophily/ anti-homophily on continuous variables).

● ***nodematch***: tendency of nodes to form ties with those of matching values (Homophily on categorical variables)

● ***nodemix***: mixing matrix of all different combinations of node attributes (ex. A -\> A ties, A-\> B ties, B -\> A ties, B -\> B ties). To avoid model overspecification, we need to leave one of these cells out of the model. The weights (effect sizes) estimated for all of the terms we leave in the model then represent the effect of a combination relative to the effect that we left out.

● ***edgecov***: covariance between edges of two networks (the presence/strength of a tie in an outside network on whether a tie exists in our dependent variable network -- Advice)
